---
title: "Results_Peishan"
output: html_document
date: "2022-12-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(forcats)
library(tidyverse)
library(GGally)
library(cowplot)
```

```{r}
Stats_by_Country_1972_1996 <- read_csv("data/Stats_by_Country_1972-1996.csv", na = "NA")
Stats_by_Country_1997_2021 <- read_csv("data/Stats_by_Country_1997-2021.csv", na = "NA")
Stats_by_Income_1972_2021 <- read_csv("data/Stats_by_Income_1972-2021.csv", na = "NA")
Stats_by_Region_1972_2021 <- read_csv("data/Stats_by_Region_1972-2021.csv", na = "NA")
```

###Region data

```{r}
# Rename Columns
setnames((Stats_by_Region_1972_2021), old = c("Series Name","Series Code", "Country Name", "Country Code"), new = c("SeN", "SeC", "CoN", "CoC"))
colnames(Stats_by_Region_1972_2021) <- substring(colnames(Stats_by_Region_1972_2021), 1, 4)
```

###Income data
```{r}
# Rename Columns
setnames((Stats_by_Income_1972_2021), old = c("Series Name","Series Code", "Country Name", "Country Code"), new = c("SeN", "SeC", "CoN", "CoC"))
colnames(Stats_by_Income_1972_2021) <- substring(colnames(Stats_by_Income_1972_2021), 1, 4)
```

###Country data
```{r}
common_col_names <- intersect(names(Stats_by_Country_1972_1996), names(Stats_by_Country_1997_2021))
Stats_by_Country_1972_2021 <- merge(Stats_by_Country_1972_1996, Stats_by_Country_1997_2021, all.x = TRUE)
```

```{r}
# Rename Columns
setnames((Stats_by_Country_1972_2021), old = c("Series Name","Series Code", "Country Name", "Country Code"), new = c("SeN", "SeC", "CoN", "CoC"))
colnames(Stats_by_Country_1972_2021) <- substring(colnames(Stats_by_Country_1972_2021), 1, 4)
```

## Highlighting the current situation and historical trends. 

### Overall trends of population change/ Life expectancy/mortality rate for the whole world in the past 50 years. 

1a. Total population change in regions categorized by income level: 
```{r}
Income_Population <- Stats_by_Income_1972_2021 %>%
  filter(SeN == "Population, total") %>%
  select(-c("SeN","SeC","CoC")) 
```

```{r}
#pivot_longer
Income_population_clean <- Income_Population %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Total_Population")

Income_population_clean$Total_Population <- as.numeric(Income_population_clean$Total_Population)
Income_population_clean$Years <- as.numeric(Income_population_clean$Years)

Income_population_clean <- Income_population_clean %>%
  mutate(lag = lag(Total_Population)) %>%
  mutate(pop.change = ((Income_population_clean$Total_Population - lag) / lag)) %>%
  filter(Years != "1972") %>%
  mutate(CoN = fct_reorder2(CoN, Years, pop.change))
```

```{r}
ggplot(Income_population_clean, aes(x = Years, y = pop.change, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Total Population change in regions categorized by income level")
```
1b. Population change in regions categorized by geography: 
```{r}
Geo_Population <- Stats_by_Region_1972_2021 %>%
  filter(SeN == "Population, total") %>%
  select(-c("SeN","SeC","CoC"))
```

```{r}
#pivot_longer
Geo_population_clean <- Geo_Population %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Total_Population")

Geo_population_clean$Total_Population <- as.numeric(Geo_population_clean$Total_Population)
Geo_population_clean$Years <- as.numeric(Geo_population_clean$Years)

Geo_population_clean <- Geo_population_clean %>%
  mutate(lag = lag(Total_Population)) %>%
  mutate(pop.change = ((Geo_population_clean$Total_Population - lag) / lag)) %>%
  filter(Years != "1972") %>%
  mutate(CoN = fct_reorder2(CoN, Years, pop.change))

```

```{r}
ggplot(Geo_population_clean, aes(x = Years, y = pop.change, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Total population change in regions categorized by geography")
```
1c. Population change in regions categorized by income - gender disparity. 

female: 
```{r}
Income_fPopulation <- Stats_by_Income_1972_2021 %>%
  filter(SeN == "Population, female") %>%
  select(-c("SeN","SeC","CoC"))
```

```{r}
#pivot_longer
Income_fpopulation_clean <- Income_fPopulation %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Female_Population")

Income_fpopulation_clean$Years <- as.numeric(Income_fpopulation_clean$Years)
Income_fpopulation_clean$Female_Population <- as.numeric(Income_fpopulation_clean$Female_Population)

Income_fpopulation_clean <- Income_fpopulation_clean %>%
  mutate(lag = lag(Female_Population)) %>%
  mutate(fpop.change = ((Income_fpopulation_clean$Female_Population - lag) / lag)) %>%
  filter(Years != "1972")%>%
  mutate(CoN = fct_reorder2(CoN, Years, fpop.change))

```

```{r}
fpop_I <- 
ggplot(Income_fpopulation_clean, aes(x = Years, y = fpop.change, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Female population change in regions categorized by income")
```

male: 
```{r}
Income_mPopulation <- Stats_by_Income_1972_2021 %>%
  filter(SeN == "Population, male") %>%
  select(-c("SeN","SeC","CoC"))
```

```{r}
#pivot_longer
Income_mpopulation_clean <- Income_mPopulation %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Male_Population")

Income_mpopulation_clean$Years <- as.numeric(Income_mpopulation_clean$Years)
Income_mpopulation_clean$Male_Population <- as.numeric(Income_mpopulation_clean$Male_Population)

Income_mpopulation_clean <- Income_mpopulation_clean %>%
  mutate(lag = lag(Male_Population)) %>%
  mutate(mpop.change = ((Income_mpopulation_clean$Male_Population - lag) / lag)) %>%
  filter(Years != "1972")%>%
  mutate(CoN = fct_reorder2(CoN, Years, mpop.change))

```

```{r}
mpop_I <- 
ggplot(Income_mpopulation_clean, aes(x = Years, y = mpop.change, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Male population change in regions categorized by income")
```

```{r}
plot_grid(fpop_I, mpop_I, align = "v", nrow = 2)
```

1d. Population change in regions categorized by geography - gender disparity. 

female: 
```{r}
Geo_fPopulation <- Stats_by_Region_1972_2021 %>%
  filter(SeN == "Population, female") %>%
  select(-c("SeN","SeC","CoC"))
```

```{r}
#pivot_longer
Geo_fpopulation_clean <- Geo_fPopulation %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Female_Population")

Geo_fpopulation_clean$Years <- as.numeric(Geo_fpopulation_clean$Years)
Geo_fpopulation_clean$Female_Population <- as.numeric(Geo_fpopulation_clean$Female_Population)

Geo_fpopulation_clean <- Geo_fpopulation_clean %>%
  mutate(lag = lag(Female_Population)) %>%
  mutate(fpop.change = ((Geo_fpopulation_clean$Female_Population - lag) / lag)) %>%
  filter(Years != "1972")%>%
  mutate(CoN = fct_reorder2(CoN, Years, fpop.change))

```

```{r}
fpop_G <- 
ggplot(Geo_fpopulation_clean, aes(x = Years, y = fpop.change, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Female population change in regions categorized by geography")
```

male: 
```{r}
Geo_mPopulation <- Stats_by_Region_1972_2021 %>%
  filter(SeN == "Population, male") %>%
  select(-c("SeN","SeC","CoC"))
```

```{r}
#pivot_longer
Geo_mpopulation_clean <- Geo_mPopulation %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Male_Population")

Geo_mpopulation_clean$Years <- as.numeric(Geo_mpopulation_clean$Years)
Geo_mpopulation_clean$Male_Population <- as.numeric(Geo_mpopulation_clean$Male_Population)

Geo_mpopulation_clean <- Geo_mpopulation_clean %>%
  mutate(lag = lag(Male_Population)) %>%
  mutate(mpop.change = ((Geo_mpopulation_clean$Male_Population - lag) / lag)) %>%
  filter(Years != "1972")%>%
  mutate(CoN = fct_reorder2(CoN, Years, mpop.change))
```

```{r}
mpop_G <- 
ggplot(Geo_mpopulation_clean, aes(x = Years, y = mpop.change, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Male population change in regions categorized by geography")
```

```{r}
plot_grid(fpop_G, mpop_G, align = "v", nrow = 2)
```

2a. Life expectancy at birth in regions categorized by income level: 
```{r}
Income_Lexpect <- Stats_by_Income_1972_2021 %>%
  filter(SeN == "Life expectancy at birth, total (years)") %>%
  select(-c("SeN","SeC","CoC","2021"))
```

```{r}
#pivot_longer
Income_Lexpect_clean <- Income_Lexpect %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Life_expectancy_at_birth")

Income_Lexpect_clean$Life_expectancy_at_birth <- as.numeric(Income_Lexpect_clean$Life_expectancy_at_birth)
Income_Lexpect_clean$Years <- as.numeric(Income_Lexpect_clean$Years)

Income_Lexpect_clean <- Income_Lexpect_clean %>%
  mutate(CoN = fct_reorder2(CoN, Years, Life_expectancy_at_birth))
```

```{r}
ggplot(Income_Lexpect_clean, aes(x = Years, y = Life_expectancy_at_birth, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Life expectancy at birth in regions categorized by income level")
```
for high income and OECD members, life expectancy dropped a little these years, why? 

2b. Life expectancy at birth in regions categorized by geography: 
```{r}
Geo_Lexpect <- Stats_by_Region_1972_2021 %>%
  filter(SeN == "Life expectancy at birth, total (years)") %>%
  select(-c("SeN","SeC","CoC","2021"))
```

```{r}
#pivot_longer
Geo_Lexpect_clean <- Geo_Lexpect %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Life_expectancy_at_birth")

Geo_Lexpect_clean$Life_expectancy_at_birth <- as.numeric(Geo_Lexpect_clean$Life_expectancy_at_birth)
Geo_Lexpect_clean$Years <- as.numeric(Geo_Lexpect_clean$Years)

Geo_Lexpect_clean <- Geo_Lexpect_clean %>%
  mutate(CoN = fct_reorder2(CoN, Years, Life_expectancy_at_birth))

```

```{r}
ggplot(Geo_Lexpect_clean, aes(x = Years, y = Life_expectancy_at_birth, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Life expectancy at birth in regions categorized by geography")
```
3a. Crude Birth rate in regions categorized by income level: 
```{r}
Income_RBirth <- Stats_by_Income_1972_2021 %>%
  filter(SeN == "Birth rate, crude (per 1,000 people)") %>%
  select(-c("SeN","SeC","CoC","2021"))
```

```{r}
#pivot_longer
Income_RBirth_clean <- Income_RBirth %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Birth_per_1000")

Income_RBirth_clean$Birth_per_1000 <- as.numeric(Income_RBirth_clean$Birth_per_1000)
Income_RBirth_clean$Years <- as.numeric(Income_RBirth_clean$Years)

Income_RBirth_clean <- Income_RBirth_clean %>%
  mutate(CoN = fct_reorder2(CoN, Years, Birth_per_1000))

```

```{r}
ggplot(Income_RBirth_clean, aes(x = Years, y = Birth_per_1000, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Birth rate crude (per 1000 people) in regions categorized by income level")
```

3b. Crude Birth rate in regions categorized by geography: 
```{r}
Geo_RBirth <- Stats_by_Region_1972_2021 %>%
  filter(SeN == "Birth rate, crude (per 1,000 people)") %>%
  select(-c("SeN","SeC","CoC","2021"))
```

```{r}
#pivot_longer
Geo_RBirth_clean <- Geo_RBirth %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Birth_per_1000")
Geo_RBirth_clean$Birth_per_1000 <- as.numeric(Geo_RBirth_clean$Birth_per_1000)
Geo_RBirth_clean$Years <- as.numeric(Geo_RBirth_clean$Years)
Geo_RBirth_clean <- Geo_RBirth_clean %>%
  mutate(CoN = fct_reorder2(CoN, Years, Birth_per_1000))
```

```{r}
ggplot(Geo_RBirth_clean, aes(x = Years, y = Birth_per_1000, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Birth rate crude (per 1000 people) in regions categorized by geography")
```


4a. Crude Death rate in regions categorized by income level: 
```{r}
Income_RDeath <- Stats_by_Income_1972_2021 %>%
  filter(SeN == "Death rate, crude (per 1,000 people)") %>%
  select(-c("SeN","SeC","CoC","2021"))
```

```{r}
#pivot_longer
Income_RDeath_clean <- Income_RDeath %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Death_per_1000")

Income_RDeath_clean$Death_per_1000 <- as.numeric(Income_RDeath_clean$Death_per_1000)
Income_RDeath_clean$Years <- as.numeric(Income_RDeath_clean$Years)

Income_RDeath_clean <- Income_RDeath_clean %>%
  mutate(CoN = fct_reorder2(CoN, Years, Death_per_1000))
```

```{r}
ggplot(Income_RDeath_clean, aes(x = Years, y = Death_per_1000, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Death rate crude (per 1000 people) in regions categorized by income level")
```

4b. Crude Death rate in regions categorized by geography: 
```{r}
Geo_RDeath <- Stats_by_Region_1972_2021 %>%
  filter(SeN == "Death rate, crude (per 1,000 people)") %>%
  select(-c("SeN","SeC","CoC","2021"))
```

```{r}
#pivot_longer
Geo_RDeath_clean <- Geo_RDeath %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Death_per_1000")
Geo_RDeath_clean$Death_per_1000 <- as.numeric(Geo_RDeath_clean$Death_per_1000)
Geo_RDeath_clean$Years <- as.numeric(Geo_RDeath_clean$Years)

```

```{r}
ggplot(Geo_RDeath_clean, aes(x = Years, y = Death_per_1000, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Death rate crude (per 1000 people) in regions categorized by geography")
```

5a. longevity by gender(income)
female: 
```{r}
Income_Rfsurv <- Stats_by_Income_1972_2021 %>%
  filter(SeN == "Survival to age 65, female (% of cohort)") %>%
  select(-c("SeN","SeC","CoC","2021"))
```

```{r}
#pivot_longer
Income_Rfsurv_clean <- Income_Rfsurv %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Survival_to_age_65_female")
Income_Rfsurv_clean$Survival_to_age_65_female <- as.numeric(Income_Rfsurv_clean$Survival_to_age_65_female)
Income_Rfsurv_clean$Years <- as.numeric(Income_RDeath_clean$Years)
```

```{r}
Rfsurv_I <- 
  ggplot(Income_Rfsurv_clean, aes(x = Years, y = Survival_to_age_65_female, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Percentage of female survived to 65 in regions categorized by income level")
```

male: 
```{r}
Income_Rmsurv <- Stats_by_Income_1972_2021 %>%
  filter(SeN == "Survival to age 65, male (% of cohort)") %>%
  select(-c("SeN","SeC","CoC","2021"))
```

```{r}
#pivot_longer
Income_Rmsurv_clean <- Income_Rmsurv %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Survival_to_age_65_male")
Income_Rmsurv_clean$Survival_to_age_65_male <- as.numeric(Income_Rmsurv_clean$Survival_to_age_65_male)
Income_Rmsurv_clean$Years <- as.numeric(Income_Rmsurv_clean$Years)
```

```{r}
Rmsurv_I <- 
ggplot(Income_Rmsurv_clean, aes(x = Years, y = Survival_to_age_65_male, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Percentage of male survived to 65 in regions categorized by income level")
```

```{r}
#Rmfsurv <- list(Rfsurv, Rmsurv)
plot_grid(Rfsurv_I, Rmsurv_I, align = "v", nrow = 2)
```

5b. longevity by gender(region)
female: 
```{r}
Geo_Rfsurv <- Stats_by_Region_1972_2021 %>%
  filter(SeN == "Survival to age 65, female (% of cohort)") %>%
  select(-c("SeN","SeC","CoC","2021"))
```

```{r}
#pivot_longer
Geo_Rfsurv_clean <- Geo_Rfsurv %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Survival_to_age_65_female")
Geo_Rfsurv_clean$Survival_to_age_65_female <- as.numeric(Geo_Rfsurv_clean$Survival_to_age_65_female)
Geo_Rfsurv_clean$Years <- as.numeric(Geo_RDeath_clean$Years)
```

```{r}
Rfsurv_G <- 
  ggplot(Geo_Rfsurv_clean, aes(x = Years, y = Survival_to_age_65_female, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Percentage of female survived to 65 in regions categorized by geography")
```

male: 
```{r}
Geo_Rmsurv <- Stats_by_Region_1972_2021 %>%
  filter(SeN == "Survival to age 65, male (% of cohort)") %>%
  select(-c("SeN","SeC","CoC","2021"))
```

```{r}
#pivot_longer
Geo_Rmsurv_clean <- Geo_Rmsurv %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Survival_to_age_65_male")
Geo_Rmsurv_clean$Survival_to_age_65_male <- as.numeric(Geo_Rmsurv_clean$Survival_to_age_65_male)
Geo_Rmsurv_clean$Years <- as.numeric(Geo_Rmsurv_clean$Years)
```

```{r}
Rmsurv_G <- 
ggplot(Geo_Rmsurv_clean, aes(x = Years, y = Survival_to_age_65_male, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Percentage of male survived to 65 in regions categorized by geography")
```

```{r}
plot_grid(Rfsurv_G, Rmsurv_G, align = "v", nrow = 2)
```

## Looking for patterns and relationships among demographic and other factors
* How do other demographic factors affect birth rate and population?




