---
title: "results2_Peishan"
output: html_document
date: "2022-12-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(forcats)
library(tidyverse)
library(GGally)
library(cowplot)
library(data.table)
library(plotly)
library(ggridges)
```

```{r}
Stats_by_Country_1972_1996 <- read_csv("data/Stats_by_Country_1972-1996.csv", na = "NA")
Stats_by_Country_1997_2021 <- read_csv("data/Stats_by_Country_1997-2021.csv", na = "NA")
Stats_by_Income_1972_2021 <- read_csv("data/Stats_by_Income_1972-2021.csv", na = "NA")
Stats_by_Region_1972_2021 <- read_csv("data/Stats_by_Region_1972-2021.csv", na = "NA")
```

###Region data

```{r}
# Rename Columns
setnames((Stats_by_Region_1972_2021), old = c("Series Name","Series Code", "Country Name", "Country Code"), new = c("SeN", "SeC", "CoN", "CoC"))
colnames(Stats_by_Region_1972_2021) <- substring(colnames(Stats_by_Region_1972_2021), 1, 4)
```

###Income data
```{r}
# Rename Columns
setnames((Stats_by_Income_1972_2021), old = c("Series Name","Series Code", "Country Name", "Country Code"), new = c("SeN", "SeC", "CoN", "CoC"))
colnames(Stats_by_Income_1972_2021) <- substring(colnames(Stats_by_Income_1972_2021), 1, 4)
```

###Country data
```{r}
common_col_names <- intersect(names(Stats_by_Country_1972_1996), names(Stats_by_Country_1997_2021))
Stats_by_Country_1972_2021 <- merge(Stats_by_Country_1972_1996, Stats_by_Country_1997_2021, all.x = TRUE)
```

```{r}
# Rename Columns
setnames((Stats_by_Country_1972_2021), old = c("Series Name","Series Code", "Country Name", "Country Code"), new = c("SeN", "SeC", "CoN", "CoC"))
colnames(Stats_by_Country_1972_2021) <- substring(colnames(Stats_by_Country_1972_2021), 1, 4)
```

##################################
##########  Question 1  ##########
##################################

**Line chart**
a. Total population in regions categorized by income level: 
```{r}
Income_Population <- Stats_by_Income_1972_2021 %>%
  filter(SeN == "Population, total") %>%
  select(-c("SeN","SeC","CoC")) 
```

```{r}
#pivot_longer
Income_population_clean <- Income_Population %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Total_Population")

Income_population_clean$Total_Population <- as.numeric(Income_population_clean$Total_Population)
Income_population_clean$Years <- as.numeric(Income_population_clean$Years)

Income_population_clean <- Income_population_clean %>%
  mutate(CoN = fct_reorder2(CoN, Years, Total_Population))
```

```{r}
ggplot(Income_population_clean, aes(x = Years, y = Total_Population, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Total Population in regions categorized by income level")
```
b. Total population in regions categorized by geographical info: 
```{r}
Geo_Population <- Stats_by_Region_1972_2021 %>%
  filter(SeN == "Population, total") %>%
  select(-c("SeN","SeC","CoC")) 
```

```{r}
#pivot_longer
Geo_population_clean <- Geo_Population %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Total_Population")

Geo_population_clean$Total_Population <- as.numeric(Geo_population_clean$Total_Population)
Geo_population_clean$Years <- as.numeric(Geo_population_clean$Years)

Geo_population_clean <- Geo_population_clean %>%
  mutate(CoN = fct_reorder2(CoN, Years, Total_Population))
```

```{r}
ggplot(Geo_population_clean, aes(x = Years, y = Total_Population, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Total Population in regions categorized by geography")
```

**Percent stacked bar plot**

```{r}
ggplot(Income_population_clean, aes(x = Years, y = Total_Population, fill = CoN)) +
  geom_bar(position = "fill",stat = "identity") +
  ggtitle("Change of the relative frequency of population in different regions categorized by income with time")
```


```{r}
ggplot(Geo_population_clean, aes(x = Years, y = Total_Population, fill = CoN)) +
  geom_bar(position = "fill",stat = "identity") +
  ggtitle("Change of the relative frequency of population in different regions categorized by geography with time")
```





##################################
##########  Question 3  ##########
##################################

Cleveland dot plot (1980):  
x = Birth rate x 10, Death rate x 10, Population growth (annual %)  
y = region
  
Cleveland dot plot (2020):  
Same

**Cleveland dot plot**

For regions categorized by income level in 1980: 
```{r}
bdp_1980 <- Stats_by_Income_1972_2021 %>%
  select ("SeN", "CoN", "1980") %>%
  filter(SeN == "Birth rate, crude (per 1,000 people)" | SeN == "Death rate, crude (per 1,000 people)" | SeN == "Population growth (annual %)") %>%
  select(-"SeN")
bdp_1980_b <- bdp_1980[1:7,]
bdp_1980_d <- bdp_1980[8:14,]
bdp_1980_g <- bdp_1980[15:21,]
bdp_1980_clean <- cbind(bdp_1980_b, bdp_1980_d, bdp_1980_g)
bdp_1980_clean <- bdp_1980_clean[!duplicated(as.list(bdp_1980_clean))]
setnames(bdp_1980_clean, old = c("1980","1980.1", "1980.2"), new = c("rbirth", "rdeath", "gpop"))
bdp_1980_clean$rbirth <- as.numeric(bdp_1980_clean$rbirth)/10
bdp_1980_clean$rdeath <- as.numeric(bdp_1980_clean$rdeath)/10
bdp_1980_clean$gpop <- as.numeric(bdp_1980_clean$gpop)/1
```

```{r}
bdp_1980_clean %>%
  gather(key='measures',value='value',rbirth, rdeath, gpop) %>%
  ggplot(aes(x=value, y=fct_reorder2(CoN, measures == "gpop",value,.desc=FALSE), color = measures)) +
  geom_point() +
  ggtitle("1980's Birth rate & Death rate & Population Growth in regions categorized by income") +
  ylab("") +
  theme_linedraw()
```
For regions categorized by geography in 1980: 
```{r}
bdpg_1980 <- Stats_by_Region_1972_2021 %>%
  select ("SeN", "CoN", "1980") %>%
  filter(SeN == "Birth rate, crude (per 1,000 people)" | SeN == "Death rate, crude (per 1,000 people)" | SeN == "Population growth (annual %)") %>%
  select(-"SeN")
```

```{r}
bdpg_1980_b <- bdpg_1980[1:9,]
bdpg_1980_d <- bdpg_1980[10:18,]
bdpg_1980_g <- bdpg_1980[19:27,]
bdpg_1980_clean <- cbind(bdpg_1980_b, bdpg_1980_d, bdpg_1980_g)
bdpg_1980_clean <- bdpg_1980_clean[!duplicated(as.list(bdpg_1980_clean))]
setnames(bdpg_1980_clean, old = c("1980","1980.1", "1980.2"), new = c("rbirth", "rdeath", "gpop"))
bdpg_1980_clean$rbirth <- as.numeric(bdpg_1980_clean$rbirth)/10
bdpg_1980_clean$rdeath <- as.numeric(bdpg_1980_clean$rdeath)/10
bdpg_1980_clean$gpop <- as.numeric(bdpg_1980_clean$gpop)/1
```

```{r}
bdpg_1980_clean %>%
  gather(key='measures',value='value',rbirth, rdeath, gpop) %>%
  ggplot(aes(x=value, y=fct_reorder2(CoN, measures == "gpop",value,.desc=FALSE), color = measures)) +
  geom_point() +
  ggtitle("1980's Birth rate & Death rate & Population Growth in regions categorized by geography") +
  ylab("") +
  theme_linedraw()
```
For regions categorized by income level in 2020: 
```{r}
bdp_2020 <- Stats_by_Income_1972_2021 %>%
  select ("SeN", "CoN", "2020") %>%
  filter(SeN == "Birth rate, crude (per 1,000 people)" | SeN == "Death rate, crude (per 1,000 people)" | SeN == "Population growth (annual %)") %>%
  select(-"SeN")
bdp_2020_b <- bdp_2020[1:7,]
bdp_2020_d <- bdp_2020[8:14,]
bdp_2020_g <- bdp_2020[15:21,]
bdp_2020_clean <- cbind(bdp_2020_b, bdp_2020_d, bdp_2020_g)
bdp_2020_clean <- bdp_2020_clean[!duplicated(as.list(bdp_2020_clean))]
setnames(bdp_2020_clean, old = c("2020","2020.1", "2020.2"), new = c("rbirth", "rdeath", "gpop"))
bdp_2020_clean$rbirth <- as.numeric(bdp_2020_clean$rbirth)/10
bdp_2020_clean$rdeath <- as.numeric(bdp_2020_clean$rdeath)/10
bdp_2020_clean$gpop <- as.numeric(bdp_2020_clean$gpop)/1
```

```{r}
bdp_2020_clean %>%
  gather(key='measures',value='value',rbirth, rdeath, gpop) %>%
  ggplot(aes(x=value, y=fct_reorder2(CoN, measures == "gpop",value,.desc=FALSE), color = measures)) +
  geom_point() +
  ggtitle("2020's Birth rate & Death rate & Population Growth in regions categorized by income") +
  ylab("") +
  theme_linedraw()
```
For regions categorized by geography in 2020: 
```{r}
bdpg_2020 <- Stats_by_Region_1972_2021 %>%
  select ("SeN", "CoN", "2020") %>%
  filter(SeN == "Birth rate, crude (per 1,000 people)" | SeN == "Death rate, crude (per 1,000 people)" | SeN == "Population growth (annual %)") %>%
  select(-"SeN")
```

```{r}
bdpg_2020_b <- bdpg_2020[1:9,]
bdpg_2020_d <- bdpg_2020[10:18,]
bdpg_2020_g <- bdpg_2020[19:27,]
bdpg_2020_clean <- cbind(bdpg_2020_b, bdpg_2020_d, bdpg_2020_g)
bdpg_2020_clean <- bdpg_2020_clean[!duplicated(as.list(bdpg_2020_clean))]
setnames(bdpg_2020_clean, old = c("2020","2020.1", "2020.2"), new = c("rbirth", "rdeath", "gpop"))
bdpg_2020_clean$rbirth <- as.numeric(bdpg_2020_clean$rbirth)/10
bdpg_2020_clean$rdeath <- as.numeric(bdpg_2020_clean$rdeath)/10
bdpg_2020_clean$gpop <- as.numeric(bdpg_2020_clean$gpop)/1
```

```{r}
bdpg_2020_clean %>%
  gather(key='measures',value='value',rbirth, rdeath, gpop) %>%
  ggplot(aes(x=value, y=fct_reorder2(CoN, measures == "gpop",value,.desc=FALSE), color = measures)) +
  geom_point() +
  ggtitle("2020's Birth rate & Death rate & Population Growth in regions categorized by geography") +
  ylab("") +
  theme_linedraw()
```
##################################
##########  Question 4  ##########
##################################

Scatter plot:  
x = birth rate by country in 1980  
y = birth rate by country in 2020

```{r}
br_8020 <- Stats_by_Country_1972_2021 %>%
  select ("SeN", "CoN", "1980", "2020") %>%
  filter(SeN == "Birth rate, crude (per 1,000 people)")
pt_8020 <- Stats_by_Country_1972_2021 %>%
  select ("SeN", "CoN", "1980", "2020") %>%
  filter(SeN == "Population, total")
brpt_8020 <- cbind(br_8020, pt_8020)
brpt_8020 <- brpt_8020[!duplicated(as.list(brpt_8020))]
setnames(brpt_8020, old = c("1980","2020", "1980.1","2020.1"), new = c("br1980", "br2020","pop1980","pop2020"))
brpt_8020 <- brpt_8020 %>%
  filter(as.numeric(pop1980) > 1000000 & as.numeric(pop2020) > 1000000) %>%
  select(-c("SeN","SeN.1", "pop1980", "pop2020"))
brpt_8020$br1980 <- as.numeric(brpt_8020$br1980)
brpt_8020$br2020 <- as.numeric(brpt_8020$br2020)
```

```{r}
brpt_8020_p <- brpt_8020 %>%
  ggplot(aes(br1980, br2020, color = CoN))+
  geom_point()+
  theme_bw()
ggplotly(brpt_8020_p)
```

##################################
##########  Question 5  ##########
##################################

Ridgeline plot (by country):  
x = Life expectancy at birth, total (years)  
y = 1980, 1990, 2000, 2010, 2020

```{r}
Lexpect <- Stats_by_Country_1972_2021 %>%
  select ("SeN", "CoN", "1980", "1990","2000", "2010", "2020") %>%
  filter(SeN == "Life expectancy at birth, total (years)")
Tpop <- Stats_by_Country_1972_2021 %>%
  select ("SeN", "CoN", "1980", "1990","2000", "2010", "2020") %>%
  filter(SeN == "Population, total")
L_T <- cbind(Lexpect, Tpop)
L_T <- L_T[!duplicated(as.list(L_T))]
setnames(L_T, old = c("1980.1","1990.1","2000.1","2010.1","2020.1"), new = c("Pop1980","Pop1990","Pop2000","Pop2010","Pop2020"))
L_T <- L_T %>%
  filter(as.numeric(Pop1980) > 1000000 & as.numeric(Pop1990) > 1000000 & as.numeric(Pop2000) > 1000000 & as.numeric(Pop2010) > 1000000 & as.numeric(Pop2020) > 1000000) %>%
  select(-c("SeN","SeN.1", "Pop1980","Pop1990","Pop2000","Pop2010","Pop2020"))
L_T$"1980" <- as.numeric(L_T$"1980")
L_T$"1990" <- as.numeric(L_T$"1990")
L_T$"2000" <- as.numeric(L_T$"2000")
L_T$"2010" <- as.numeric(L_T$"2010")
L_T$"2020" <- as.numeric(L_T$"2020")
```

```{r}
L_T_clean <- L_T %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Life_expectancy_at_birth")
L_T_clean$Life_expectancy_at_birth <- as.numeric(L_T_clean$Life_expectancy_at_birth)
```

```{r}
ggplot(L_T_clean, aes(x = Life_expectancy_at_birth, y = Years)) +
  geom_density_ridges(fill = "blue",alpha = .5, scale = 1.2) +
  ggtitle("World longevity over the past decades") +
  ylab("Years")
```

##################################
##########  Question 9  ##########
##################################

```{r}
Income_Lexpect <- Stats_by_Income_1972_2021 %>%
  filter(SeN == "Life expectancy at birth, total (years)") %>%
  select(-c("SeN","SeC","CoC","2021"))
```

```{r}
#pivot_longer
Income_Lexpect_clean <- Income_Lexpect %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Life_expectancy_at_birth")

Income_Lexpect_clean$Life_expectancy_at_birth <- as.numeric(Income_Lexpect_clean$Life_expectancy_at_birth)
Income_Lexpect_clean$Years <- as.numeric(Income_Lexpect_clean$Years)

Income_Lexpect_clean <- Income_Lexpect_clean %>%
  mutate(CoN = fct_reorder2(CoN, Years, Life_expectancy_at_birth))
```

```{r}
ggplot(Income_Lexpect_clean, aes(x = Years, y = Life_expectancy_at_birth, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Life expectancy at birth in regions categorized by income level")
```

##################################
##########  Question 10 ##########
##################################

Heat map (or mosaic plot) (income table, 2020):  
vertical cut = income category  
horizontal cut =  
Cause of death, by communicable diseases and maternal, prenatal and nutrition conditions (% of total)  
Cause of death, by injury (% of total)  
Cause of death, by non-communicable diseases (% of total)

```{r}
Income_Cdeath <- Stats_by_Income_1972_2021 %>%
  select ("SeN", "CoN", "2019") %>%
  filter(SeN == "Cause of death, by communicable diseases and maternal, prenatal and nutrition conditions (% of total)" | SeN == "Cause of death, by injury (% of total)" | SeN == "Cause of death, by non-communicable diseases (% of total)")
Income_Cdeath$"2019"<- as.numeric(Income_Cdeath$"2019")
Income_Cdeath$SeN <- replace(Income_Cdeath$SeN, Income_Cdeath$SeN == "Cause of death, by communicable diseases and maternal, prenatal and nutrition conditions (% of total)","communicable")
Income_Cdeath$SeN <- replace(Income_Cdeath$SeN, Income_Cdeath$SeN == "Cause of death, by injury (% of total)","injury")
Income_Cdeath$SeN <- replace(Income_Cdeath$SeN, Income_Cdeath$SeN == "Cause of death, by non-communicable diseases (% of total)","non-communicable")
```

```{r}
ggplot(Income_Cdeath, aes(x = CoN, y = SeN)) +
  geom_tile(aes(fill = Income_Cdeath$"2019"), color = "white") +
  coord_fixed()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  ggtitle("2019's causes of death in different income categories")
```

##################################
##########  Question 11 ##########
##################################

Parallel coordinate plot (by country):  
2020 divided by 2000:  
GNI per capita growth 

2000:  
Labor force total / Population total  
Urban population (% of total population)  
Literacy rate, adult total (% of people ages 15 and above)  
School enrollment, secondary (% gross)  
School enrollment, tertiary (% gross)  

```{r}
GNI_0020 <- Stats_by_Country_1972_2021 %>%
  select ("SeN", "CoN", "2000", "2020") %>%
  filter(SeN == "GNI per capita, Atlas method (current US$)")
Tpop_0020 <- Stats_by_Country_1972_2021 %>%
  select ("SeN","2000") %>%
  filter(SeN == "Population, total")
Lforce_0020 <- Stats_by_Country_1972_2021 %>%
  select ("SeN","2000") %>%
  filter(SeN == "Labor force, total")
Upop_0020 <- Stats_by_Country_1972_2021 %>%
  select ("SeN","2000") %>%
  filter(SeN == "Urban population (% of total population)")
Lrat_0020 <- Stats_by_Country_1972_2021 %>%
  select ("SeN","2000") %>%
  filter(SeN == "Literacy rate, adult total (% of people ages 15 and above)")
Ses_0020 <- Stats_by_Country_1972_2021 %>%
  select ("SeN","2000") %>%
  filter(SeN == "School enrollment, secondary (% gross)")
Set_0020 <- Stats_by_Country_1972_2021 %>%
  select ("SeN","2000") %>%
  filter(SeN == "School enrollment, tertiary (% gross)")
```

a. Labor force total / Population total vs. GNI Growth
```{r}
GNI_Tpop <- cbind(GNI_0020,Tpop_0020,Lforce_0020)
GNI_Tpop <- GNI_Tpop[!duplicated(as.list(GNI_Tpop))]
GNI_Tpop[GNI_Tpop == '..'] <- NA
GNI_Tpop <- GNI_Tpop %>% 
  drop_na()
GNI_Tpop$"2000" <- as.numeric(GNI_Tpop$"2000")
GNI_Tpop$"2020" <- as.numeric(GNI_Tpop$"2020")
GNI_Tpop$"2000.1" <- as.numeric(GNI_Tpop$"2000.1")
GNI_Tpop$"2000.2" <- as.numeric(GNI_Tpop$"2000.2")
GNI_Tpop <- GNI_Tpop %>% 
  mutate(GNI_Growth = (GNI_Tpop$"2020"-GNI_Tpop$"2000"/GNI_Tpop$"2000")) %>% 
  mutate(Labor_Pop = (GNI_Tpop$"2000.2"/GNI_Tpop$"2000.1")) %>%
  select("CoN","GNI_Growth","Labor_Pop")
```

```{r}
ggparcoord(GNI_Tpop, columns = 3:2, alphaLines = .3)+
  ggtitle("Labor force total / Population total vs. GNI")
```

b. Urban population (% of total population) vs. GNI Growth
```{r}
GNI_Upop <- cbind(GNI_0020,Upop_0020)
GNI_Upop <- GNI_Upop[!duplicated(as.list(GNI_Upop))]
GNI_Upop[GNI_Upop == '..'] <- NA
GNI_Upop <- GNI_Upop %>% 
  drop_na()
GNI_Upop$"2000" <- as.numeric(GNI_Upop$"2000")
GNI_Upop$"2020" <- as.numeric(GNI_Upop$"2020")
GNI_Upop$"2000.1" <- as.numeric(GNI_Upop$"2000.1")
GNI_Upop <- GNI_Upop %>% 
  mutate(GNI_Growth = (GNI_Upop$"2020"-GNI_Upop$"2000"/GNI_Upop$"2000")) %>% 
  mutate(Urban_Pop = (GNI_Upop$"2000.1")) %>%
  select("CoN","GNI_Growth","Urban_Pop")
```

```{r}
ggparcoord(GNI_Upop, columns = 3:2, alphaLines = .3)+
  ggtitle("Urban population (% of total population) vs. GNI Growth")
```

c. Literacy rate, adult total (% of people ages 15 and above)  vs. GNI Growth

```{r}
GNI_Lrat <- cbind(GNI_0020,Lrat_0020)
GNI_Lrat <- GNI_Lrat[!duplicated(as.list(GNI_Lrat))]
GNI_Lrat[GNI_Lrat == '..'] <- NA
GNI_Lrat <- GNI_Lrat %>% 
  drop_na()
GNI_Lrat$"2000" <- as.numeric(GNI_Lrat$"2000")
GNI_Lrat$"2020" <- as.numeric(GNI_Lrat$"2020")
GNI_Lrat$"2000.1" <- as.numeric(GNI_Lrat$"2000.1")
GNI_Lrat <- GNI_Lrat %>% 
  mutate(GNI_Growth = (GNI_Lrat$"2020"-GNI_Lrat$"2000"/GNI_Lrat$"2000")) %>% 
  mutate(Literacy_adult = (GNI_Lrat$"2000.1")) %>%
  select("CoN","GNI_Growth","Literacy_adult")
```

```{r}
ggparcoord(GNI_Lrat, columns = 3:2, alphaLines = .3)+
  ggtitle("Literacy rate, adult total (% of people ages 15 and above) vs. GNI Growth")
```

d. School enrollment, secondary (% gross)  vs. GNI Growth

```{r}
GNI_Ses <- cbind(GNI_0020,Ses_0020)
GNI_Ses <- GNI_Ses[!duplicated(as.list(GNI_Ses))]
GNI_Ses[GNI_Ses == '..'] <- NA
GNI_Ses <- GNI_Ses %>% 
  drop_na()
GNI_Ses$"2000" <- as.numeric(GNI_Ses$"2000")
GNI_Ses$"2020" <- as.numeric(GNI_Ses$"2020")
GNI_Ses$"2000.1" <- as.numeric(GNI_Ses$"2000.1")
GNI_Ses <- GNI_Ses %>% 
  mutate(GNI_Growth = (GNI_Ses$"2020"-GNI_Ses$"2000"/GNI_Ses$"2000")) %>% 
  mutate(Secondary_enrollment = (GNI_Ses$"2000.1")) %>%
  select("CoN","GNI_Growth","Secondary_enrollment")
```

```{r}
ggparcoord(GNI_Ses, columns = 3:2, alphaLines = .3)+
  ggtitle("School enrollment, secondary (% gross) vs. GNI Growth")
```

e. School enrollment, tertiary (% gross) vs. GNI Growth

```{r}
GNI_Set <- cbind(GNI_0020,Set_0020)
GNI_Set <- GNI_Set[!duplicated(as.list(GNI_Set))]
GNI_Set[GNI_Set == '..'] <- NA
GNI_Set <- GNI_Set %>% 
  drop_na()
GNI_Set$"2000" <- as.numeric(GNI_Set$"2000")
GNI_Set$"2020" <- as.numeric(GNI_Set$"2020")
GNI_Set$"2000.1" <- as.numeric(GNI_Set$"2000.1")
GNI_Set <- GNI_Set %>% 
  mutate(GNI_Growth = (GNI_Set$"2020"-GNI_Set$"2000"/GNI_Set$"2000")) %>% 
  mutate(Tertiary_enrollment = (GNI_Set$"2000.1")) %>%
  select("CoN","GNI_Growth","Tertiary_enrollment")
```

```{r}
ggparcoord(GNI_Set, columns = 3:2, alphaLines = .3)+
  ggtitle("School enrollment, tertiary (% gross) vs. GNI Growth")
```

