---
title: "results2_Peishan"
output: html_document
date: "2022-12-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(forcats)
library(tidyverse)
library(GGally)
library(cowplot)
library(data.table)
library(plotly)
```

```{r}
Stats_by_Country_1972_1996 <- read_csv("data/Stats_by_Country_1972-1996.csv", na = "NA")
Stats_by_Country_1997_2021 <- read_csv("data/Stats_by_Country_1997-2021.csv", na = "NA")
Stats_by_Income_1972_2021 <- read_csv("data/Stats_by_Income_1972-2021.csv", na = "NA")
Stats_by_Region_1972_2021 <- read_csv("data/Stats_by_Region_1972-2021.csv", na = "NA")
```

###Region data

```{r}
# Rename Columns
setnames((Stats_by_Region_1972_2021), old = c("Series Name","Series Code", "Country Name", "Country Code"), new = c("SeN", "SeC", "CoN", "CoC"))
colnames(Stats_by_Region_1972_2021) <- substring(colnames(Stats_by_Region_1972_2021), 1, 4)
```

###Income data
```{r}
# Rename Columns
setnames((Stats_by_Income_1972_2021), old = c("Series Name","Series Code", "Country Name", "Country Code"), new = c("SeN", "SeC", "CoN", "CoC"))
colnames(Stats_by_Income_1972_2021) <- substring(colnames(Stats_by_Income_1972_2021), 1, 4)
```

###Country data
```{r}
common_col_names <- intersect(names(Stats_by_Country_1972_1996), names(Stats_by_Country_1997_2021))
Stats_by_Country_1972_2021 <- merge(Stats_by_Country_1972_1996, Stats_by_Country_1997_2021, all.x = TRUE)
```

```{r}
# Rename Columns
setnames((Stats_by_Country_1972_2021), old = c("Series Name","Series Code", "Country Name", "Country Code"), new = c("SeN", "SeC", "CoN", "CoC"))
colnames(Stats_by_Country_1972_2021) <- substring(colnames(Stats_by_Country_1972_2021), 1, 4)
```

##################################
##########  Question 1  ##########
##################################

**Line chart**
a. Total population in regions categorized by income level: 
```{r}
Income_Population <- Stats_by_Income_1972_2021 %>%
  filter(SeN == "Population, total") %>%
  select(-c("SeN","SeC","CoC")) 
```

```{r}
#pivot_longer
Income_population_clean <- Income_Population %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Total_Population")

Income_population_clean$Total_Population <- as.numeric(Income_population_clean$Total_Population)
Income_population_clean$Years <- as.numeric(Income_population_clean$Years)

Income_population_clean <- Income_population_clean %>%
  mutate(CoN = fct_reorder2(CoN, Years, Total_Population))
```

```{r}
ggplot(Income_population_clean, aes(x = Years, y = Total_Population, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Total Population in regions categorized by income level")
```
b. Total population in regions categorized by geographical info: 
```{r}
Geo_Population <- Stats_by_Region_1972_2021 %>%
  filter(SeN == "Population, total") %>%
  select(-c("SeN","SeC","CoC")) 
```

```{r}
#pivot_longer
Geo_population_clean <- Geo_Population %>%
  pivot_longer(cols = !CoN, names_to = "Years", values_to = "Total_Population")

Geo_population_clean$Total_Population <- as.numeric(Geo_population_clean$Total_Population)
Geo_population_clean$Years <- as.numeric(Geo_population_clean$Years)

Geo_population_clean <- Geo_population_clean %>%
  mutate(CoN = fct_reorder2(CoN, Years, Total_Population))
```

```{r}
ggplot(Geo_population_clean, aes(x = Years, y = Total_Population, color = CoN))+
  geom_line()+
  geom_point()+
  ggtitle("Total Population in regions categorized by geography")
```

**Mosaic chart**

##################################
##########  Question 3  ##########
##################################

Cleveland dot plot (1980):  
x = Birth rate x 10, Death rate x 10, Population growth (annual %)  
y = region
  
Cleveland dot plot (2020):  
Same

**Cleveland dot plot**

For regions categorized by income level in 1980: 
```{r}
bdp_1980 <- Stats_by_Income_1972_2021 %>%
  select ("SeN", "CoN", "1980") %>%
  filter(SeN == "Birth rate, crude (per 1,000 people)" | SeN == "Death rate, crude (per 1,000 people)" | SeN == "Population growth (annual %)") %>%
  select(-"SeN")
bdp_1980_b <- bdp_1980[1:7,]
bdp_1980_d <- bdp_1980[8:14,]
bdp_1980_g <- bdp_1980[15:21,]
bdp_1980_clean <- cbind(bdp_1980_b, bdp_1980_d, bdp_1980_g)
bdp_1980_clean <- bdp_1980_clean[!duplicated(as.list(bdp_1980_clean))]
setnames(bdp_1980_clean, old = c("1980","1980.1", "1980.2"), new = c("rbirth", "rdeath", "gpop"))
bdp_1980_clean$rbirth <- as.numeric(bdp_1980_clean$rbirth)/10
bdp_1980_clean$rdeath <- as.numeric(bdp_1980_clean$rdeath)/10
bdp_1980_clean$gpop <- as.numeric(bdp_1980_clean$gpop)/1
```

```{r}
bdp_1980_clean %>%
  gather(key='measures',value='value',rbirth, rdeath, gpop) %>%
  ggplot(aes(x=value, y=fct_reorder2(CoN, measures == "gpop",value,.desc=FALSE), color = measures)) +
  geom_point() +
  ggtitle("1980's Birth rate & Death rate & Population Growth in regions categorized by income") +
  ylab("") +
  theme_linedraw()
```
For regions categorized by geography in 1980: 
```{r}
bdpg_1980 <- Stats_by_Region_1972_2021 %>%
  select ("SeN", "CoN", "1980") %>%
  filter(SeN == "Birth rate, crude (per 1,000 people)" | SeN == "Death rate, crude (per 1,000 people)" | SeN == "Population growth (annual %)") %>%
  select(-"SeN")
```

```{r}
bdpg_1980_b <- bdpg_1980[1:9,]
bdpg_1980_d <- bdpg_1980[10:18,]
bdpg_1980_g <- bdpg_1980[19:27,]
bdpg_1980_clean <- cbind(bdpg_1980_b, bdpg_1980_d, bdpg_1980_g)
bdpg_1980_clean <- bdpg_1980_clean[!duplicated(as.list(bdpg_1980_clean))]
setnames(bdpg_1980_clean, old = c("1980","1980.1", "1980.2"), new = c("rbirth", "rdeath", "gpop"))
bdpg_1980_clean$rbirth <- as.numeric(bdpg_1980_clean$rbirth)/10
bdpg_1980_clean$rdeath <- as.numeric(bdpg_1980_clean$rdeath)/10
bdpg_1980_clean$gpop <- as.numeric(bdpg_1980_clean$gpop)/1
```

```{r}
bdpg_1980_clean %>%
  gather(key='measures',value='value',rbirth, rdeath, gpop) %>%
  ggplot(aes(x=value, y=fct_reorder2(CoN, measures == "gpop",value,.desc=FALSE), color = measures)) +
  geom_point() +
  ggtitle("1980's Birth rate & Death rate & Population Growth in regions categorized by geography") +
  ylab("") +
  theme_linedraw()
```
For regions categorized by income level in 2020: 
```{r}
bdp_2020 <- Stats_by_Income_1972_2021 %>%
  select ("SeN", "CoN", "2020") %>%
  filter(SeN == "Birth rate, crude (per 1,000 people)" | SeN == "Death rate, crude (per 1,000 people)" | SeN == "Population growth (annual %)") %>%
  select(-"SeN")
bdp_2020_b <- bdp_2020[1:7,]
bdp_2020_d <- bdp_2020[8:14,]
bdp_2020_g <- bdp_2020[15:21,]
bdp_2020_clean <- cbind(bdp_2020_b, bdp_2020_d, bdp_2020_g)
bdp_2020_clean <- bdp_2020_clean[!duplicated(as.list(bdp_2020_clean))]
setnames(bdp_2020_clean, old = c("2020","2020.1", "2020.2"), new = c("rbirth", "rdeath", "gpop"))
bdp_2020_clean$rbirth <- as.numeric(bdp_2020_clean$rbirth)/10
bdp_2020_clean$rdeath <- as.numeric(bdp_2020_clean$rdeath)/10
bdp_2020_clean$gpop <- as.numeric(bdp_2020_clean$gpop)/1
```

```{r}
bdp_2020_clean %>%
  gather(key='measures',value='value',rbirth, rdeath, gpop) %>%
  ggplot(aes(x=value, y=fct_reorder2(CoN, measures == "gpop",value,.desc=FALSE), color = measures)) +
  geom_point() +
  ggtitle("2020's Birth rate & Death rate & Population Growth in regions categorized by income") +
  ylab("") +
  theme_linedraw()
```
For regions categorized by geography in 2020: 
```{r}
bdpg_2020 <- Stats_by_Region_1972_2021 %>%
  select ("SeN", "CoN", "2020") %>%
  filter(SeN == "Birth rate, crude (per 1,000 people)" | SeN == "Death rate, crude (per 1,000 people)" | SeN == "Population growth (annual %)") %>%
  select(-"SeN")
```

```{r}
bdpg_2020_b <- bdpg_2020[1:9,]
bdpg_2020_d <- bdpg_2020[10:18,]
bdpg_2020_g <- bdpg_2020[19:27,]
bdpg_2020_clean <- cbind(bdpg_2020_b, bdpg_2020_d, bdpg_2020_g)
bdpg_2020_clean <- bdpg_2020_clean[!duplicated(as.list(bdpg_2020_clean))]
setnames(bdpg_2020_clean, old = c("2020","2020.1", "2020.2"), new = c("rbirth", "rdeath", "gpop"))
bdpg_2020_clean$rbirth <- as.numeric(bdpg_2020_clean$rbirth)/10
bdpg_2020_clean$rdeath <- as.numeric(bdpg_2020_clean$rdeath)/10
bdpg_2020_clean$gpop <- as.numeric(bdpg_2020_clean$gpop)/1
```

```{r}
bdpg_2020_clean %>%
  gather(key='measures',value='value',rbirth, rdeath, gpop) %>%
  ggplot(aes(x=value, y=fct_reorder2(CoN, measures == "gpop",value,.desc=FALSE), color = measures)) +
  geom_point() +
  ggtitle("2020's Birth rate & Death rate & Population Growth in regions categorized by geography") +
  ylab("") +
  theme_linedraw()
```
##################################
##########  Question 4  ##########
##################################

Scatter plot:  
x = birth rate by country in 1980  
y = birth rate by country in 2020

```{r}
br_8020 <- Stats_by_Country_1972_2021 %>%
  select ("SeN", "CoN", "1980", "2020") %>%
  filter(SeN == "Birth rate, crude (per 1,000 people)")
pt_8020 <- Stats_by_Country_1972_2021 %>%
  select ("SeN", "CoN", "1980", "2020") %>%
  filter(SeN == "Population, total")
brpt_8020 <- cbind(br_8020, pt_8020)
brpt_8020 <- brpt_8020[!duplicated(as.list(brpt_8020))]
setnames(brpt_8020, old = c("1980","2020", "1980.1","2020.1"), new = c("br1980", "br2020","pop1980","pop2020"))
brpt_8020 <- brpt_8020 %>%
  filter(as.numeric(pop1980) > 1000000 & as.numeric(pop2020) > 1000000) %>%
  select(-c("SeN","SeN.1", "pop1980", "pop2020"))
brpt_8020$br1980 <- as.numeric(brpt_8020$br1980)
brpt_8020$br2020 <- as.numeric(brpt_8020$br2020)
```


```{r}
brpt_8020_p <- brpt_8020 %>%
  ggplot(aes(br1980, br2020, color = CoN))+
  geom_point()+
  theme_bw()
ggplotly(brpt_8020_p)
```

##################################
##########  Question 5  ##########
##################################
Ridgeline plot (by country):  
x = Life expectancy at birth, total (years)  
y = 1980, 1990, 2000, 2010, 2020


##################################
##########  Question 9  ##########
##################################

Ridgeline plots x 5 (1980, 1990, 2000, 2010, 2020) (income table):  
x = Life expectancy at birth, total (years)  
y = income category [income table]

##################################
##########  Question 10 ##########
##################################

Heat map (or mosaic plot) (income table, 2020):  
vertical cut = income category  
horizontal cut =  
Cause of death, by communicable diseases and maternal, prenatal and nutrition conditions (% of total)  
Cause of death, by injury (% of total)  
Cause of death, by non-communicable diseases (% of total)

##################################
##########  Question 11 ##########
##################################

Parallel coordinate plot (by country):  
2020 divided by 2000:  
GNI per capita growth 